---
title: "R Final Assignment Notebook"
output: html_notebook
---

##Data Inspection
For data inspection, I first copied the files from the [BCB546X repository](https://github.com/EEOB-BioData/BCB546X-Fall2017) to my [Assignment repository](https://github.com/dormilon/R_final_assignment) and inspected the folder from R:
```{r}
file.info(dir("Original_files/", full.names = T))
```
Here, we can see that there are 4 files on the folder and that `fang_et_al_genotypes.txt` is 11051939 bytes and `snp_position.txt` is 82763 bytes.


I second open and read the `fang_et_al_genotypes.txt` and `snp_position.txt` files using:
```{r}
genotypes <- read.table(file = "Original_files/fang_et_al_genotypes.txt", sep = "\t", header = T)
snp <- read.table(file = "Original_files/snp_position.txt", sep = "\t", header = T)
```


After creating the objects `genotypes` and `snp` I inspected their properties and dimensions using:
```{r}
str(genotypes)
str(snp)
dim(genotypes)
dim(snp)
```
From this we can appreciate that `genotypes` is a data frame object with 2,782 observations (rows) and 986 variables (columns). The `snp` object is a data frame with 983 observations (rows) and 15 variables (columns).

##Data Processing
###Prepared files for 'join':
Extracted maize- and teosinte-specific data, respectively.
```{r}
maize_genotypes <- subset(genotypes, Group == "ZMMIL" | Group == "ZMMLR" | Group == "ZMMMR")
teo_genotypes <- subset(genotypes, Group == "ZMPBA" | Group == "ZMPIL" | Group == "ZMPJA")
```

Next, I transposed the both data frames, mantaining the object type:
```{r}
transp_maize_genotypes <- as.data.frame(t(maize_genotypes))
transp_teo_genotypes <- as.data.frame(t(teo_genotypes))
```

Get rid of the first three rows on the transposed data and create a new column with the row names (this is going to be our common column for merging the data).
```{r}
transp_maize_genotypes <- transp_maize_genotypes[-c(1:3),]
transp_teo_genotypes <- transp_teo_genotypes[-c(1:3),]
transp_maize_genotypes$SNP_ID <- rownames(transp_maize_genotypes)
transp_teo_genotypes$SNP_ID <- rownames(transp_teo_genotypes)
```

Next, sorted the data frames, just to be sure:
```{r}
transp_maize_genotypes <- transp_maize_genotypes[order(transp_maize_genotypes$SNP_ID),]
transp_teo_genotypes <- transp_teo_genotypes[order(transp_teo_genotypes$SNP_ID),]
snp <- snp[order(snp$SNP_ID),]
```


###Joining the files and preparing for the extraction of the data:
One more format for the `snp` data frame before merging. I created a subset of `snp` with the requested columns:
```{r}
snp_sub <- snp[,c(1,3,4)]
```

Used `merge` to join the data:
```{r}
maize_snp_genotype <- merge(snp_sub, transp_maize_genotypes, by = "SNP_ID")
teo_snp_genotype <- merge(snp_sub, transp_teo_genotypes, by = "SNP_ID")
```

